---
title: "Data Analysis"
subtitle: 'STAT 385, Spring 2019'
author: "Ajay Dugar (dugar3)"
date: 'May 9th, 2019'
output:
  html_document:
    theme: readable
    toc: yes
    toc_float:
      collapse: false
always_allow_html: yes # required for PDF
---


```{r,echo=FALSE}
pkg_list = c("plyr", "tidyr", "dplyr", "rvest", "stringr", "ballr", "magrittr", "rsconnect", "plotly", "data.table")
mia_pkgs = pkg_list[!(pkg_list %in% installed.packages()[,"Package"])]
if(length(mia_pkgs) > 0) install.packages(mia_pkgs)
loaded_pkgs = lapply(pkg_list, require, character.only=TRUE)
```


```{r}
years = (1989:2019)
#Combining raw and advanced statistics
for (i in years){
  
  assign(paste0("players_", i), cbind(NBAPerGameAdvStatistics(season = i), NBAPerGameStatistics(season = i)))

}

for (i in years){
#remove duplicate columns
  assign(paste0("players_", i), get(paste0("players_", i))[, !duplicated(colnames(get(paste0("players_", i))))])
  
  #removing the players that are duplicated
  assign(paste0("players_", i),  get(paste0("players_", i))[!duplicated(get(paste0("players_", i))$rk), ])
  
  #Removing players who haven't played a minimum of 70% of the games in the season
  assign(paste0("players_", i),  get(paste0("players_", i))[get(paste0("players_", i))$g > 0.7*max(get(paste0("players_", i))$g),])
}

for (i in years) {
  df_name = paste0("players_",i)
  set(get(df_name), j = "player", value = str_remove_all(get(df_name)$player, "\\*"))
}
```

```{r}

year_names = c()

for (i in years){
  
    year_names = c(year_names, paste0("players_", i))
    
}

```


```{r}

for (i in year_names){
  
    assign(i, subset(get(i), select = -c(x, x_2) ))
    
}

for (i in year_names){
  
    assign(i, subset(get(i), select = -c(tm, link) ))
    
}

```


```{r}
# NBA Award Winners
players_1989 <- within(players_1989, mvp <- ifelse(players_1989$player == "Magic Johnson", 1, 0)) 
players_1989 <- within(players_1989, roy <- ifelse(players_1989$player == "Mitch Richmond", 1, 0)) 
players_1989 <- within(players_1989, dpoy <- ifelse(players_1989$player == "Mark Eaton", 1, 0)) 
players_1989 <- within(players_1989, smoy <- ifelse(players_1989$player == "Eddie Johnson", 1, 0)) 
players_1989 <- within(players_1989, mip <- ifelse(players_1989$player == "Kevin Johnson", 1, 0)) 
players_1990 <- within(players_1990, mvp <- ifelse(players_1990$player == "Magic Johnson", 1, 0)) 
players_1990 <- within(players_1990, roy <- ifelse(players_1990$player == "David Robinson", 1, 0)) 
players_1990 <- within(players_1990, dpoy <- ifelse(players_1990$player == "Dennis Rodman", 1, 0)) 
players_1990 <- within(players_1990, smoy <- ifelse(players_1990$player == "Ricky Pierce", 1, 0)) 
players_1990 <- within(players_1990, mip <- ifelse(players_1990$player == "Rony Seikaly", 1, 0))
players_1991 <- within(players_1991, mvp <- ifelse(players_1991$player == "Michael Jordan", 1, 0)) 
players_1991 <- within(players_1991, roy <- ifelse(players_1991$player == "Derrick Coleman", 1, 0)) 
players_1991 <- within(players_1991, dpoy <- ifelse(players_1991$player == "Dennis Rodman", 1, 0)) 
players_1991 <- within(players_1991, smoy <- ifelse(players_1991$player == "Detlef Schrempf", 1, 0)) 
players_1991 <- within(players_1991, mip <- ifelse(players_1991$player == "Scott Skiles", 1, 0))
players_1992 <- within(players_1992, mvp <- ifelse(players_1992$player == "Michael Jordan", 1, 0)) 
players_1992 <- within(players_1992, roy <- ifelse(players_1992$player == "Larry Johnson", 1, 0)) 
players_1992 <- within(players_1992, dpoy <- ifelse(players_1992$player == "David Robinson", 1, 0)) 
players_1992 <- within(players_1992, smoy <- ifelse(players_1992$player == "Detlef Schrempf", 1, 0)) 
players_1992 <- within(players_1992, mip <- ifelse(players_1992$player == "Pervis Ellison", 1, 0))
players_1993 <- within(players_1993, mvp <- ifelse(players_1993$player == "Charles Barkley", 1, 0)) 
players_1993 <- within(players_1993, roy <- ifelse(players_1993$player == "Shaquille O'Neal", 1, 0)) 
players_1993 <- within(players_1993, dpoy <- ifelse(players_1993$player == "Hakeem Olajuwon", 1, 0)) 
players_1993 <- within(players_1993, smoy <- ifelse(players_1993$player == "Clifford Robinson", 1, 0)) 
players_1993 <- within(players_1993, mip <- ifelse(players_1993$player == "Mahmoud Abdul-Rauf", 1, 0))
players_1994 <- within(players_1994, mvp <- ifelse(players_1994$player == "Hakeem Olajuwon", 1, 0)) 
players_1994 <- within(players_1994, roy <- ifelse(players_1994$player == "Chris Webber", 1, 0)) 
players_1994 <- within(players_1994, dpoy <- ifelse(players_1994$player == "Hakeem Olajuwon", 1, 0)) 
players_1994 <- within(players_1994, smoy <- ifelse(players_1994$player == "Dell Curry", 1, 0)) 
players_1994 <- within(players_1994, mip <- ifelse(players_1994$player == "Don MacLean", 1, 0))
players_1995 <- within(players_1995, mvp <- ifelse(players_1995$player == "David Robinson", 1, 0)) 
players_1995 <- within(players_1995, roy <- ifelse(players_1995$player == "Grant Hill" | players_1995$player == "Jason Kidd", 1, 0)) 
players_1995 <- within(players_1995, dpoy <- ifelse(players_1995$player == "Dikembe Mutombo", 1, 0)) 
players_1995 <- within(players_1995, smoy <- ifelse(players_1995$player == "Anthony Mason", 1, 0)) 
players_1995 <- within(players_1995, mip <- ifelse(players_1995$player == "Dana Barros", 1, 0))
players_1996 <- within(players_1996, mvp <- ifelse(players_1996$player == "Michael Jordan", 1, 0)) 
players_1996 <- within(players_1996, roy <- ifelse(players_1996$player == "Damon Stoudamire", 1, 0)) 
players_1996 <- within(players_1996, dpoy <- ifelse(players_1996$player == "Gary Payton", 1, 0)) 
players_1996 <- within(players_1996, smoy <- ifelse(players_1996$player == "Toni Kukoc", 1, 0)) 
players_1996 <- within(players_1996, mip <- ifelse(players_1996$player == "Gheorghe Muresan", 1, 0))
players_1997 <- within(players_1997, mvp <- ifelse(players_1997$player == "Karl Malone", 1, 0)) 
players_1997 <- within(players_1997, roy <- ifelse(players_1997$player == "Allen Iverson", 1, 0)) 
players_1997 <- within(players_1997, dpoy <- ifelse(players_1997$player == "Dikembe Mutombo", 1, 0)) 
players_1997 <- within(players_1997, smoy <- ifelse(players_1997$player == "John Starks", 1, 0)) 
players_1997 <- within(players_1997, mip <- ifelse(players_1997$player == "Isaac Austin", 1, 0))
players_1998 <- within(players_1998, mvp <- ifelse(players_1998$player == "Michael Jordon", 1, 0)) 
players_1998 <- within(players_1998, roy <- ifelse(players_1998$player == "Tim Duncan", 1, 0)) 
players_1998 <- within(players_1998, dpoy <- ifelse(players_1998$player == "Dikembe Mutombo", 1, 0)) 
players_1998 <- within(players_1998, smoy <- ifelse(players_1998$player == "Danny Manning", 1, 0)) 
players_1998 <- within(players_1998, mip <- ifelse(players_1998$player == "Alan Henderson", 1, 0))
players_1999 <- within(players_1999, mvp <- ifelse(players_1999$player == "Karl Malone", 1, 0)) 
players_1999 <- within(players_1999, roy <- ifelse(players_1999$player == "Vince Carter", 1, 0)) 
players_1999 <- within(players_1999, dpoy <- ifelse(players_1999$player == "Alonzo Mourning", 1, 0)) 
players_1999 <- within(players_1999, smoy <- ifelse(players_1999$player == "Darrell Armstrong", 1, 0)) 
players_1999 <- within(players_1999, mip <- ifelse(players_1999$player == "Darrell Armstrong", 1, 0))
players_2000 <- within(players_2000, mvp <- ifelse(players_2000$player == "Shaquille O'Neal", 1, 0)) 
players_2000 <- within(players_2000, roy <- ifelse(players_2000$player == "Elton Brand" | players_2000$player == "Steve Francis", 1, 0)) 
players_2000 <- within(players_2000, dpoy <- ifelse(players_2000$player == "Alonzo Mourning", 1, 0)) 
players_2000 <- within(players_2000, smoy <- ifelse(players_2000$player == "Rodney Rogers", 1, 0)) 
players_2000 <- within(players_2000, mip <- ifelse(players_2000$player == "Jalen Rose", 1, 0))
players_2001 <- within(players_2001, mvp <- ifelse(players_2001$player == "Allen Iverson", 1, 0)) 
players_2001 <- within(players_2001, roy <- ifelse(players_2001$player == "Mike Miller", 1, 0)) 
players_2001 <- within(players_2001, dpoy <- ifelse(players_2001$player == "Dikembe Mutombo", 1, 0)) 
players_2001 <- within(players_2001, smoy <- ifelse(players_2001$player == "Aaron McKie", 1, 0)) 
players_2001 <- within(players_2001, mip <- ifelse(players_2001$player == "Tracy McGrady", 1, 0))
players_2002 <- within(players_2002, mvp <- ifelse(players_2002$player == "Tim Duncan", 1, 0)) 
players_2002 <- within(players_2002, roy <- ifelse(players_2002$player == "Pau Gasol", 1, 0)) 
players_2002 <- within(players_2002, dpoy <- ifelse(players_2002$player == "Ben Wallace", 1, 0)) 
players_2002 <- within(players_2002, smoy <- ifelse(players_2002$player == "Corliss Williamson", 1, 0)) 
players_2002 <- within(players_2002, mip <- ifelse(players_2002$player == "Jermaine O'Neal", 1, 0))
players_2003 <- within(players_2003, mvp <- ifelse(players_2003$player == "Tim Duncan", 1, 0)) 
players_2003 <- within(players_2003, roy <- ifelse(players_2003$player == "Amar'e Stoudemire", 1, 0)) 
players_2003 <- within(players_2003, dpoy <- ifelse(players_2003$player == "Ben Wallace", 1, 0)) 
players_2003 <- within(players_2003, smoy <- ifelse(players_2003$player == "Bobby Jackson", 1, 0)) 
players_2003 <- within(players_2003, mip <- ifelse(players_2003$player == "Gilbert Arenas", 1, 0))
players_2004 <- within(players_2004, mvp <- ifelse(players_2004$player == "Kevin Garnett", 1, 0)) 
players_2004 <- within(players_2004, roy <- ifelse(players_2004$player == "LeBron James", 1, 0)) 
players_2004 <- within(players_2004, dpoy <- ifelse(players_2004$player == "Metta World Peace", 1, 0)) 
players_2004 <- within(players_2004, smoy <- ifelse(players_2004$player == "Antawn Jamison", 1, 0)) 
players_2004 <- within(players_2004, mip <- ifelse(players_2004$player == "Zach Randolph", 1, 0))
players_2005 <- within(players_2005, mvp <- ifelse(players_2005$player == "Steve Nash", 1, 0)) 
players_2005 <- within(players_2005, roy <- ifelse(players_2005$player == "Emeka Okafor", 1, 0)) 
players_2005 <- within(players_2005, dpoy <- ifelse(players_2005$player == "Ben Wallace", 1, 0)) 
players_2005 <- within(players_2005, smoy <- ifelse(players_2005$player == "Ben Gordon", 1, 0)) 
players_2005 <- within(players_2005, mip <- ifelse(players_2005$player == "Bobby Simmons", 1, 0))
players_2006 <- within(players_2006, mvp <- ifelse(players_2006$player == "Steve Nash", 1, 0)) 
players_2006 <- within(players_2006, roy <- ifelse(players_2006$player == "Chris Paul", 1, 0)) 
players_2006 <- within(players_2006, dpoy <- ifelse(players_2006$player == "Ben Wallace", 1, 0)) 
players_2006 <- within(players_2006, smoy <- ifelse(players_2006$player == "Mike Miller", 1, 0)) 
players_2006 <- within(players_2006, mip <- ifelse(players_2006$player == "Boris Diaw", 1, 0))
players_2007 <- within(players_2007, mvp <- ifelse(players_2007$player == "Dirk Nowitzki", 1, 0)) 
players_2007 <- within(players_2007, roy <- ifelse(players_2007$player == "Brandon Roy", 1, 0)) 
players_2007 <- within(players_2007, dpoy <- ifelse(players_2007$player == "Marcus Camby", 1, 0)) 
players_2007 <- within(players_2007, smoy <- ifelse(players_2007$player == "Leandro Barbosa", 1, 0)) 
players_2007 <- within(players_2007, mip <- ifelse(players_2007$player == "Monta Ellis", 1, 0))
players_2008 <- within(players_2008, mvp <- ifelse(players_2008$player == "Kobe Bryant", 1, 0)) 
players_2008 <- within(players_2008, roy <- ifelse(players_2008$player == "Kevin Durant", 1, 0)) 
players_2008 <- within(players_2008, dpoy <- ifelse(players_2008$player == "Kevin Garnett", 1, 0)) 
players_2008 <- within(players_2008, smoy <- ifelse(players_2008$player == "Manu Ginobili", 1, 0)) 
players_2008 <- within(players_2008, mip <- ifelse(players_2008$player == "Hedo Turkoglu", 1, 0))
players_2009 <- within(players_2009, mvp <- ifelse(players_2009$player == "LeBron James", 1, 0)) 
players_2009 <- within(players_2009, roy <- ifelse(players_2009$player == "Derrick Rose", 1, 0)) 
players_2009 <- within(players_2009, dpoy <- ifelse(players_2009$player == "Dwight Howard", 1, 0)) 
players_2009 <- within(players_2009, smoy <- ifelse(players_2009$player == "Jason Terry", 1, 0)) 
players_2009 <- within(players_2009, mip <- ifelse(players_2009$player == "Danny Granger", 1, 0))
players_2010 <- within(players_2010, mvp <- ifelse(players_2010$player == "LeBron James", 1, 0)) 
players_2010 <- within(players_2010, roy <- ifelse(players_2010$player == "Tyreke Evans", 1, 0)) 
players_2010 <- within(players_2010, dpoy <- ifelse(players_2010$player == "Dwight Howard", 1, 0)) 
players_2010 <- within(players_2010, smoy <- ifelse(players_2010$player == "Jamal Crawford", 1, 0)) 
players_2010 <- within(players_2010, mip <- ifelse(players_2010$player == "Aaron Brooks", 1, 0))
players_2011 <- within(players_2011, mvp <- ifelse(players_2011$player == "Derrick Rose", 1, 0)) 
players_2011 <- within(players_2011, roy <- ifelse(players_2011$player == "Blake Griffin", 1, 0)) 
players_2011 <- within(players_2011, dpoy <- ifelse(players_2011$player == "Dwight Howard", 1, 0)) 
players_2011 <- within(players_2011, smoy <- ifelse(players_2011$player == "Lamar Odom", 1, 0)) 
players_2011 <- within(players_2011, mip <- ifelse(players_2011$player == "Kevin Love", 1, 0))
players_2012 <- within(players_2012, mvp <- ifelse(players_2012$player == "LeBron James", 1, 0)) 
players_2012 <- within(players_2012, roy <- ifelse(players_2012$player == "Kyrie Irving", 1, 0)) 
players_2012 <- within(players_2012, dpoy <- ifelse(players_2012$player == "Tyson Chandler", 1, 0)) 
players_2012 <- within(players_2012, smoy <- ifelse(players_2012$player == "James Harden", 1, 0)) 
players_2012 <- within(players_2012, mip <- ifelse(players_2012$player == "Ryan Anderson", 1, 0))
players_2013 <- within(players_2013, mvp <- ifelse(players_2013$player == "LeBron James", 1, 0)) 
players_2013 <- within(players_2013, roy <- ifelse(players_2013$player == "Damian Lillard", 1, 0)) 
players_2013 <- within(players_2013, dpoy <- ifelse(players_2013$player == "Tyson Chandler", 1, 0)) 
players_2013 <- within(players_2013, smoy <- ifelse(players_2013$player == "Marc Gasol", 1, 0)) 
players_2013 <- within(players_2013, mip <- ifelse(players_2013$player == "J.R. Smith", 1, 0))
players_2014 <- within(players_2014, mvp <- ifelse(players_2014$player == "Kevin Durant", 1, 0)) 
players_2014 <- within(players_2014, roy <- ifelse(players_2014$player == "Michael Carter-Williams", 1, 0)) 
players_2014 <- within(players_2014, dpoy <- ifelse(players_2014$player == "Joakim Noah", 1, 0)) 
players_2014 <- within(players_2014, smoy <- ifelse(players_2014$player == "Jamal Crawford", 1, 0)) 
players_2014 <- within(players_2014, mip <- ifelse(players_2014$player == "Goran Dragic", 1, 0))
players_2015 <- within(players_2015, mvp <- ifelse(players_2015$player == "Stephen Curry", 1, 0)) 
players_2015 <- within(players_2015, roy <- ifelse(players_2015$player == "Andrew Wiggins", 1, 0)) 
players_2015 <- within(players_2015, dpoy <- ifelse(players_2015$player == "Kawhi Leonard", 1, 0)) 
players_2015 <- within(players_2015, smoy <- ifelse(players_2015$player == "Lou Williams", 1, 0)) 
players_2015 <- within(players_2015, mip <- ifelse(players_2015$player == "Jimmy Butler", 1, 0))
players_2016 <- within(players_2016, mvp <- ifelse(players_2016$player == "Stephen Curry", 1, 0)) 
players_2016 <- within(players_2016, roy <- ifelse(players_2016$player == "Karl-Anthony Towns", 1, 0)) 
players_2016 <- within(players_2016, dpoy <- ifelse(players_2016$player == "Kawhi Leonard", 1, 0)) 
players_2016 <- within(players_2016, smoy <- ifelse(players_2016$player == "Jamal Crawford", 1, 0)) 
players_2016 <- within(players_2016, mip <- ifelse(players_2016$player == "CJ McCollum", 1, 0))
players_2017 <- within(players_2017, mvp <- ifelse(players_2017$player == "Russell Westbrook", 1, 0)) 
players_2017 <- within(players_2017, roy <- ifelse(players_2017$player == "Malcolm Brogdon", 1, 0)) 
players_2017 <- within(players_2017, dpoy <- ifelse(players_2017$player == "Draymond Green", 1, 0)) 
players_2017 <- within(players_2017, smoy <- ifelse(players_2017$player == "Eric Gordon", 1, 0)) 
players_2017 <- within(players_2017, mip <- ifelse(players_2017$player == "Giannis Antetokounmpo", 1, 0))
players_2018 <- within(players_2018, mvp <- ifelse(players_2018$player == "James Harden", 1, 0)) 
players_2018 <- within(players_2018, roy <- ifelse(players_2018$player == "Ben Simmons", 1, 0)) 
players_2018 <- within(players_2018, dpoy <- ifelse(players_2018$player == "Rudy Gobert", 1, 0)) 
players_2018 <- within(players_2018, smoy <- ifelse(players_2018$player == "Lou Williams", 1, 0)) 
players_2018 <- within(players_2018, mip <- ifelse(players_2018$player == "Victor Oladipo", 1, 0))
```


```{r}
year_names = c()
for (i in years){
  
    year_names = c(year_names, paste0("players_", i))
    
}
```


```{r}
for (i in year_names){
  
    assign(i, subset(get(i), select = -c(x, x_2) ))
    
}
for (i in year_names){
  
    assign(i, subset(get(i), select = -c(tm, link) ))
    
}
```


```{r}

#per + tspercent + x3par + ftr + orbpercent + drbpercent + trbpercent + astpercent + stlpercent + blkpercent + tovpercent + usgpercent + ows + dws + obpm + dbpm + vorp + g + mp +  gs + x3p + x3pa + x2p + x2pa + ft + fta + orb + drb + trb + ast + stl + blk + tov + pf + pts
```


```{r}
players_all = players_1989[FALSE,]
for (i in year_names){
  
  players_all <- rbind(players_all, get(i))  
  
}



```

```{r}
players_all$PG = ifelse(players_all$pos == "PG", 1, 0)
players_all$SG = ifelse(players_all$pos == "SG", 1, 0)
players_all$SF = ifelse(players_all$pos == "SF", 1, 0)
players_all$PF = ifelse(players_all$pos == "PF", 1, 0)
players_all$C = ifelse(players_all$pos == "C", 1, 0)
```



```{r}
#mvp
indep.vars_normal <- ~age + mp + x3p + x3pa + x2p + x2pa + ft + fta + orb + drb + ast + stl + blk + tov + pf + PG + SG + SF + PF + C 
indep.vars_advanced <- ~age + mp + per + tspercent + orbpercent + drbpercent + astpercent + blkpercent + tovpercent + usgpercent + ows + dws + PG + SG + SF + PF + C 
object <- glm(mvp ~ 1, family=binomial(link="logit"), data=players_all)
#swout_mvp_normal <- step(object, scope = indep.vars_normal, direction = "both")
#summary(swout_mvp_normal)
#vcov(swout_mvp_normal)

swout_mvp_advanced <- step(object, scope = indep.vars_advanced, direction = "both")
summary(swout_mvp_advanced) #better
vcov(swout_mvp_advanced)
```

```{r}
#roy
object <- glm(roy ~ 1, family=binomial(link="logit"), data=players_all)
swout_roy_normal <- step(object, scope = indep.vars_normal, direction = "both")
summary(swout_roy_normal) #better
#vcov(swout_roy_normal)
#swout_roy_advanced <- step(object, scope = indep.vars_advanced, direction = "both")
#summary(swout_roy_advanced)
#vcov(swout_roy_advanced)
```

```{r}
# dpoy
object <- glm(dpoy ~ 1, family=binomial(link="logit"), data=players_all)
#swout_dpoy_normal <- step(object, scope = indep.vars_normal, direction = "both")
#summary(swout_dpoy_normal)
#vcov(swout_dpoy_normal)

swout_dpoy_advanced <- step(object, scope = indep.vars_advanced, direction = "both")
summary(swout_dpoy_advanced) #better
#vcov(swout_dpoy_advanced)
```

```{r}
#smoy
object <- glm(smoy ~ 1, family=binomial(link="logit"), data=players_all[which(players_all$gs < 41),])
#swout_smoy_normal <- step(object, scope = indep.vars_normal, direction = "both")
#summary(swout_smoy_normal)
#vcov(swout_smoy_normal)
swout_smoy_advanced <- step(object, scope = indep.vars_advanced, direction = "both")
summary(swout_smoy_advanced) #better
#vcov(swout_smoy_advanced)
```


```{r}
# mip
object <- glm(mip ~ 1, family=binomial(link="logit"), data=players_all)
swout_mip_normal <- step(object, scope = indep.vars_normal, direction = "both")
summary(swout_mip_normal) #better
#vcov(swout_mip_normal)
#swout_mip_advanced <- step(object, scope = indep.vars_advanced, direction = "both")
#summary(swout_mip_advanced)
#vcov(swout_mip_advanced)
# # mip
# 
# object <- glm(mip ~ 1, family=binomial(link="logit"), data=players_all)
# #swout_mip_normal <- step(object, scope = indep.vars_normal, direction = "both")
# #summary(swout_mip_normal)
# 
# #vcov(swout_mip_normal)
# 
# 
# #swout_mip_advanced <- step(object, scope = indep.vars_advanced, direction = "both")
# summary(swout_mip_advanced) #better
# 
# #vcov(swout_mip_advanced)

```

```{r}
#applying the model
players_2019$PG = ifelse(players_2019$pos == "PG", 1, 0)
players_2019$SG = ifelse(players_2019$pos == "SG", 1, 0)
players_2019$SF = ifelse(players_2019$pos == "SF", 1, 0)
players_2019$PF = ifelse(players_2019$pos == "PF", 1, 0)
players_2019$C = ifelse(players_2019$pos == "C", 1, 0)
```

```{r}
players_2019$mvp = as.numeric(1/(1+exp(-1*(-29.55824 + 0.65215*players_2019$per + 0.25109*players_2019$tovpercent + 0.56602*players_2019$dws))))
players_2019$mvp_odds = players_2019$mvp/sum(players_2019$mvp)
mvp_2019 = players_2019[with(players_2019, order(-mvp_odds)),]
```

```{r}
rookies_2019 = anti_join(players_2019, players_all, by = 'player')
rookies_2019$roy = as.numeric(1/(1+exp(-1*(12.8345287 + -1.0762097*rookies_2019$age + 0.0012159*rookies_2019$mp + 0.7284919*rookies_2019$orb + 0.4127871*rookies_2019$ast))))
rookies_2019$roy_odds = rookies_2019$roy/sum(rookies_2019$roy)
players_2019 <- merge(players_2019, rookies_2019, all = TRUE)

players_2019$mvp_odds = players_2019$mvp/sum(players_2019$mvp)*100

players_2019[with(players_2019, order(-mvp_odds)),]

```

```{r}

rookies_2019 = anti_join(players_2019, players_all,  by='player')


rookies_2019$roy = as.numeric(1/(1+exp(-1*(12.8345287 + -1.0762097*rookies_2019$age + 0.0012159*rookies_2019$mp + 0.7284919*rookies_2019$orb + 0.4127871*rookies_2019$ast))))

rookies_2019$roy_odds = rookies_2019$roy/sum(rookies_2019$roy)*100

players_2019 <- join(players_2019, rookies_2019, type = "left")
players_2019[is.na(players_2019)] <- 0
roy_2019 = players_2019[with(players_2019, order(-roy_odds)),]
```


```{r}
players_2019$dpoy = as.numeric(1/(1+exp(-1*(-7.6440321 + 1.6298305*players_2019$dws + -0.3212441*players_2019$usgpercent + 0.3284959*players_2019$per + -0.0018602*players_2019$mp))))
players_2019$dpoy_odds = players_2019$dpoy/sum(players_2019$dpoy)
dpoy_2019 = players_2019[with(players_2019, order(-dpoy_odds)),]
```

```{r}
players_2019$smoy = ifelse(players_2019$gs < 41, as.numeric(1/(1+exp(-1*(-4.334e+01 + 4.575e-01*players_2019$usgpercent + 1.206e+00*players_2019$dws + -1.090e+00*players_2019$blkpercent + 4.005e+01*players_2019$tspercent + 2.762e-03*players_2019$mp)))), 0)
players_2019$smoy_odds = players_2019$smoy/sum(players_2019$smoy)
smoy_2019 = players_2019[with(players_2019, order(-smoy_odds)),]
  
```

```{r}
mip_2019_new = join(players_2019, players_2018, by = "player", type = "inner")
mip_2019_old = join(players_2018, players_2019, by = "player", type = "inner")
player = mip_2019_new$player
mip_2019 = data.frame(player)
mip_2019$age = mip_2019_new$age
mip_2019$tov = mip_2019_new$tov - mip_2019_old$tov
mip_2019$x3p = mip_2019_new$x3p - mip_2019_old$x3p
mip_2019$x3pa = mip_2019_new$x3pa - mip_2019_old$x3pa
mip_2019$mip = as.numeric(1/(1+exp(-1*(-5.4244556 + 0.8269561*mip_2019$tov + -0.2037588*mip_2019$age + 3.8753971*mip_2019$x3p + -1.3835910*mip_2019$x3pa)))^9)
mip_2019$mip_odds = mip_2019$mip/sum(mip_2019$mip)
test = mip_2019[with(mip_2019, order(-mip_odds)),]
displayed = within(test, rm(player))
radarchart(displayed[1,], maxmin = FALSE)
colors_in=c( rgb(0.2,0.5,0.5,0.4), rgb(0.8,0.2,0.5,0.4) , rgb(0.7,0.5,0.1,0.4) )
```


players_2019[with(players_2019, order(-smoy_odds)),]
```

```{r}
df_mvp <- data.frame(Player = players_2019$player, Position = players_2019$pos, Age = players_2019$age, Chance = players_2019$mvp_odds)  
df_mvp =df_mvp[with(df$mvp, order(-Chance)),]
df_mvp

df_roy <- data.frame(Player = players_2019$player, Position = players_2019$pos, Age = players_2019$age, Chance = players_2019$roy_odds)  
df_roy =df_roy[with(df_roy, order(-Chance)),]
df_roy

df_dpoy <- data.frame(Player = players_2019$player, Position = players_2019$pos, Age = players_2019$age, Chance = players_2019$dpoy_odds)  
df_dpoy =df_dpoy[with(df_dpoy, order(-Chance)),]
df_dpoy
```


```{r}


players_1990$mvp = as.numeric(1/(1+exp(-1*(-29.55824 + 0.65215*players_1990$per + 0.25109*players_1990$tovpercent + 0.56602*players_1990$dws))))
players_1990$mvp_odds = players_1990$mvp/sum(players_1990$mvp)
mvp_1990 = players_1990[with(players_1990, order(-mvp_odds)),]

```

```{r}


normalize_data <- function(x, year) {
  return((100)*(x - rep(min(x), length(x))) / (rep(max(x), length(x)) - rep(min(x), length(x))))
}

df_smoy <- data.frame(Player = players_2019$player, Position = players_2019$pos, Age = players_2019$age, Chance = players_2019$smoy_odds)  
df_smoy =df_smoy[with(df_smoy, order(-Chance)),]
df_smoy
```



```{r}
mvp_1990_graph_display = mvp_1990[,c("player","per", "tspercent", "pts", "ast", "trb")]
for (i in colnames(mvp_1990_graph_display)) {
  if (i == "player") {
    
    next;
  }
  mvp_1990_graph_display[,i] = normalize_data(mvp_1990_graph_display[,i])
}

p <- plot_ly(
    type = 'scatterpolar',
    fill = 'toself'
  ) %>%
  add_trace(
    r = as.numeric(mvp_1990_graph_display[1, c("per", "tspercent", "pts", "ast", "trb")]),
    theta = c('per','tspercent','pts', 'ast', "trb"),
    name = mvp_1990_graph_display[1, "player"]
  ) %>%
  layout(
    polar = list(
      radialaxis = list(
        visible = T,
        range = c(0, 100)
      )
    )
  )

  
  p
```
